---
title: "Getting Started with Binomial Functions"
author: "Natasha Yuanto Susetyo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Binomial Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(binomial)
```

## Binomial Distribution

Binomial distribution is one of the most famous probability distribution among the discrete random variables. 
The package `"binomial"` is a minimal implementation that contains various functions regarding the binomial distribution.

Assume that X is a random variable such that X ∼ Bin(n, p). 

## Mean and Variance

As with most random variables, it is useful to know about the various summary measures of X.
The expected value or mean of a binomial distribution is: $$np$$. 
This is the expected number of successes in n trials.

The variance is given by: $$np(1 − p)$$. 
Consequently, the standard deviation is simply the square root of the variance, that is: $$\sqrt{np(1 − p)}$$

The function `bin_mean()` will calculate the binomial distribution's mean.
```{r}
bin_mean(trials = 5, prob = 0.3) #answer = 1.5
```

The function `bin_variance()` will calculate the binomial distribution's variance.
```{r}
bin_variance(trials = 5, prob = 0.3) #answer = 1.05
```

### Mode

For $$0 < p < 1$$, the most likely number of success in n independent trials with probability p of success on each trial is m, the greater integer less than or equal to $$np + p$$: 

$$m = int(np + p)$$

where int denotes the integer part function. If $$np + p$$ is an integer, as in the case p = 0.5 and
n odd, then there are two most likely numbers, m and m − 1. Otherwise, there is a unique
mode.

The function `bin_mode()` will calculate the binomial distribution's mode.
```{r}
bin_mode(trials = 5, prob = 0.3) #answer = 1
```

### Skewness and Kurtosis

Other two additional measures are skewness and kurtosis. 
Skewness is a measure of the asymmetry of the probability distribution of a random variable about its mean. 
The skewness value can be positive or negative, or undefined. 
The skewness of a binomial random variable can be calculated as:

$$skewness = \dfrac{1 − 2p}{\sqrt{np(1 − p)}}$$

The Kurtosis (from greek kurtos, meaning “curved, arching”) is a measure of the “tailedness” of the probability distribution of a random variable. 
In a similar way to the concept of skewness, kurtosis is a descriptor of the shape of a probability distribution. 
For a binomial random variable, its kurtosis can be obtained as:

$$kurtosis = \dfrac{1 − 6p(1-p)}{np(1 − p)}$$

The function `bin_skewness()` will calculate the skewness of the binomial distribution.
```{r}
bin_skewness(trials = 5, prob = 0.3) #answer = 0.39036
```

The function `bin_kurtosis()` will calculate the kurtosis of the binomial distribution.
```{r}
bin_kurtosis(trials = 5, prob = 0.3) #answer = -0.247619
```

## Finding number of combinations

The function `bin_choose()` calculates the number of combinations in which `k` successes occur in `n` trials.

```{r}
bin_choose(n = 5, k = 2) #answer = 10
```

## The binomial probability

The function `bin_probability()` calculates probability of getting a certain number successes in the number of desired trials. 

```{r}
bin_probability(success = 2, trials = 5, prob = 0.5) #answer = 0.3125
```

## The binomial distribution

We can also calculate the probability of success for each number of successes possible given the number of trials. We will use the `bin_distribution()` function.

```{r}
dis1 <- bin_distribution(trials = 5, prob = 0.5)
dis1
```

## Plot of binomial distribution

We can use the function `plot()` to see the histogram of a binomial distribution for the probability for each number of successes.

```{r}
plot(dis1)
```

## The binomial cumulative distribution

We can find the binomial cumulative distribution where it calculates the cummulative probabilities from the previous number of successes using the function `bin_cumulative()`.
```{r}
bin2 <- bin_cumulative(trials = 5, prob = 0.5)
bin2
```

## Plot of binomial cumulative distribution

We can plot the binomial cumulative distribution graph by using the `plot()` function.
```{r}
plot(bin2)
```

## The binomial variable

The function `bin_variable()` which returns an obejct of class "binvar".
```{r}
bin3 <- bin_variable(trials = 10, prob = 0.3)
bin3
```

## Print the summary 

The `summary` function can return a list of class "summary.binvar" containing named elements:
  - trials : number of trials
  - prob : probability of success
  - mean : mean or expected value
  - variance : variance
  - mode : mode
  - skewness: skewness
  - kurtosis : kurtosis
```{r}
summary(bin3)
```

